Outliers
========================================================

## In-Class Exercise #50

Alalyze exam data. Find outliers. 

### Read data

```{r}
setwd("./data")
exams<-read.csv("exams_and_names-2.csv",header=T)
setwd("../")
```

### Using boxplot w/ interquartile range

Key is using measure which is natural to the data set e.g., quartiles on distribution. Recall:
- Box goes to 25% (Q1) and 75% (Q3)
- Whiskers go to 1.5x Q1-Q3 range
- Beyond that are outliers

```{r fig.width=7, fig.height=6}
with(exams,plot(Exam.1~Exam.2))

par(mfrow=c(1,2))
with(exams,boxplot(Exam.1)) # no outliers
with(exams,boxplot(Exam.2)) # three outliers

par(mfrow=c(1,1))
boxplot(exams$Exam.1,exams$Exam.2,col="blue",main="Exam Scores",
        names=c("exam1", "exam2"),ylab="Exam Score")
```

### Using z score

```{r}
sd1<-sd(exams$Exam.1,na.rm=T)
sd2<-sd(exams$Exam.2,na.rm=T)

exams$Exam.1.z<-(exams$Exam.1-mean(exams$Exam.1,na.rm=T))/sd1
exams$Exam.2.z<-(exams$Exam.2-mean(exams$Exam.2,na.rm=T))/sd2

# Note: default for scale is to set na.rm=T
exams$Exam.1.scale<-scale(exams$Exam.1)
exams$Exam.2.scale<-scale(exams$Exam.2)

head(sort(exams$Exam.2.z))
head(sort(exams$Exam.2.z,decreasing=T))
     
# No outliers
```

## In-Class Exercise #52

Alalyze exam data. Find outliers using IQR. **Considered better than mean/sd as they are sensitive to outliers.**

### Calculating / using IQR

```{r}
# 1. get 1st and 3rd quantiles
q1<-quantile(exams$Exam.2,0.25,na.rm=T)
q3<-quantile(exams$Exam.2,0.75,na.rm=T)

# 2. calc iqr
iqr<-q3-q1

# 3. find outliers
outliers<-exams[exams$Exam.2>=q3+(1.5*iqr) | exams$Exam.2<=q1-(1.5*iqr),]
outliers
# 3 outliers

```

### Example of why iqr better: outliers skew mean/sd

```{r}
data<-c(1,2,3,4,100)
boxplot(data)
# shows outliers

sd3<-sd(data)
mean1<-mean(data)
zscores<-(data-mean1)/sd3
zscores
# no outliers
```

### iqr not perfect. some distributions are broad

```{r}
# example: T distribution w/ low degrees of freedom. It's just the distribution.
boxplot(rt(50,df=1))
```

## Detecting outliers for multiple attributes

### In-class #52: Using regression

First use basic lm to plot regression line for the data.

```{r}
# load/clean data
setwd("./data")
exams<-read.csv("exams_and_names-2.csv",header=T)
setwd("../")
exams1<-exams[!is.na(exams[,3]),]

# fit model
fit<-lm(exams$Exam.2~exams$Exam.1)
summary(fit)

# plot results
plot(exams$Exam.1,exams$Exam.2,pch=19,xlab="Exam 1", ylab="Exam 2", xlim=c(100,200),
     ylim=c(100,200))
abline(fit)

# inspect residuals
sort(fit$residuals)
# recall that residuals are the **vertical** distance to the regression (fitted values) line. 
```

Now let's look at it scaling the dots by the size of their residual

```{r}
# load/clean data
setwd("./data")
exams<-read.csv("exams_and_names-2.csv",header=T)
setwd("../")
exams1<-exams[!is.na(exams[,3]),]

# fit model
fit2<-lm(exams1$Exam.2~exams1$Exam.1)

# plot results
plot(exams$Exam.1,exams$Exam.2,pch=19,xlab="Exam 1", ylab="Exam 2", xlim=c(100,200),
     ylim=c(100,200),cex=abs(fit2$residuals)/10)

abline(fit2)

plot(fit2) # if you tell R to plot a model, it takes you through a series of graphs.
```

### In-class #53: Using Clustering w/ Kmeans

First cluster the data

```{r}
# load data
setwd("./data")
data<-read.csv("exams_and_names-2.csv",header=T)
setwd("../")

# clean data taking only needed rows
x<-data[!is.na(data[,3]),2:3]

# plot the data
plot(x,pch=19,xlab="Exam 1", ylab="Exam 2")

# fit model
fit<-kmeans(x,5)

# plot results
points(fit$centers,pch=19,col="blue",cex=2)
points(x,col=fit$cluster,pch=19)

```

Different methods end up w/ different outliers. In this case, the lower three are considered outliers b/c they have their own cluster (whenk K is 5). The upper two are not as they fall w/in the range of the others when k=5. 